### 2021-09-29 paper/header.tex
add juliacon2021 full paper for proceedings review

% **************GENERATED FILE, DO NOT EDIT**************

\title{Designing Machine Learning Pipeline Toolkit for AutoML Surrogate Modeling Optimization}

\author[1]{Paulito P. Palmes}
\author[2]{Akihiro Kishimoto}
\author[1]{Radu Marinescu}
\author[3]{Parikshit Ram}
\author[1]{Elizabeth Daly}
\affil[1]{IBM Research Europe}
\affil[2]{IBM Research Japan}
\affil[3]{IBM Research USA}

\keywords{Julia, Optimization, AutoML, ML Pipeline,Classification, Prediction, Regression}

\hypersetup{
pdftitle = {My JuliaCon proceeding},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, Optimization, Game theory, Compiler},
}