### 2023-06-22 paper/header.tex

% **************GENERATED FILE, DO NOT EDIT**************

\title{Designing Machine Learning Pipeline Toolkit for AutoML Surrogate Modeling Optimization}

\author[1]{Paulito P. Palmes}
\author[2]{Akihiro Kishimoto}
\author[1]{Radu Marinescu}
\author[3]{Parikshit Ram}
\author[1]{Elizabeth Daly}
\affil[1]{IBM Research Europe, Dublin Research Lab}
\affil[2]{IBM Research Japan}
\affil[3]{IBM Research USA}

\keywords{Julia, Optimization, AutoMLPipeline, AutoML, ML Pipeline Search, Feature Engineering, Modeling, Classification, Prediction, Regression}

\hypersetup{
pdftitle = {Designing Machine Learning Pipeline Toolkit for AutoML Surrogate Modeling Optimization},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Paulito P. Palmes, Akihiro Kishimoto, Radu Marinescu, Parikshit Ram, Elizabeth Daly},
pdfkeywords = {Julia, Optimization, AutoMLPipeline, AutoML, ML Pipeline Search, Feature Engineering, Modeling, Classification, Prediction, Regression},
}